---
title: "Simple unstratified Beale"
author: "Steve Westenbroek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simple unstratified Beale}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Data preparation

In order to use the beale function and supporting classes, a dataframe must be prepared with the following fields:

    1) *date* the sample date, encoded as an R `Date` type,
    2) *discharge* the daily mean discharge, provided in cubic meters per second, and
    3) *concentration* the concentration corresponding to the daily discharge, in milligrams per liter (NA if no sample taken that day).

```{r, initialization, warning=FALSE,message=FALSE}

library(autobeale)
library(dplyr)


```

The `autobeale::beale` function expects the incoming data to be provided in pairwise fashion. There should be a discharge value provided for *every* day within the timeframe for the laoding calculation, with a concentration value provided on days that were sampled. `dplyr::left_join` can be used to quickly create an input dataset suitable for use with the `beale` function.

```{r, results='asis'}

discharge_file     <- system.file( "extdata", "ROCKFLOW.DAT", package="autobeale")
concentration_file <- system.file( "extdata", "ROCKNO23.DAT", package="autobeale")

discharge_data     <- read.table( discharge_file, header=TRUE )
concentration_data <- read.table( concentration_file, header=TRUE )
discharge_conc_df  <- NULL
discharge_conc_df  <- dplyr::left_join(discharge_data, concentration_data, by=c("date" = "date") )

# don't forget to convert the discharge value to cubic meters per second!
discharge_conc_df$discharge <- cfs_to_cms( discharge_conc_df$discharge )

```

The resulting dataframe looks like this:

```{r, echo=FALSE, results='asis'}
knitr::kable(head(discharge_conc_df, 10))
```

## Example #1: load by means of the unstratified Beale ratio estimator

Once the dataset is in good shape, we can call `beale` to calculate the load. Note that the `beale` function expects discharge to be provided in units of *cubic meters per second*, with concentration data provided in units of *milligrams per liter*.

```{r, unstratified_beale, results='asis'}
result_list <- beale( discharge_cms=discharge_conc_df$discharge, concentration_mg_L=discharge_conc_df$concentration )
print( result_list, digits=2 )
```


