---
title: "Using autobeale support classes"
author: "Steve Westenbroek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{autobeale support classes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, results='asis'}

library(autobeale)

discharge_file     <- system.file( "extdata", "ROCKFLOW.DAT", package="autobeale")
concentration_file <- system.file( "extdata", "ROCKNO23.DAT", package="autobeale")

discharge_data     <- read.table( discharge_file, header=TRUE )
concentration_data <- read.table( concentration_file, header=TRUE )
qcdata <- dplyr::left_join(discharge_data, concentration_data, by=c("date" = "date") )

qcdata$discharge <- cfs_to_cms( qcdata$discharge )

```

## Example #2: Beale Ratio Estimator with data partitioned into five stratums

In addition to the `beale` function are a number of `R6` classes aimed at making it easier to examine various data partitioning strategies. Strategies may be ranked by their impact on the root mean square of error associated with the load estimate. In this example, a genetic algorithm package is wired up to assist in the search for an optimum partitioning strategy, with optimimum being defined as minimization of the root mean square of error associated with the overall loading estimate, similar to the original AutoBeale.

In order to use these classes, real R date classes (POSIX) must be supplied along with the discharge and concentration data. Another of Hadley Wickham's library packages, `lubridate`, makes this easy:

```{r, date munging, message=FALSE }

library(autobeale)
library(lubridate)
library(mcga)

# we're modifying the 'qcdata' dataframe from example #1
qcdata$date <- lubridate::ymd( qcdata$date )

```

The `qcdata` dataframe now looks like this:

```{r, echo=FALSE, results='asis'}
knitr::kable(head(qcdata, 10))
```

## Use of the support classes `strata` and `partitions`

First we creata a `strata` object of the appropriate size (5 stratums):

```{r, strata and partitioning}
number_of_stratums <- 5
mystrata <- strata$new( number_of_stratums, qcdata )
```

Next we wire up a genetic algoritm package to test alternative partitioning strategies. Package `mcga` uses its `evalFunc` argument to pass a vector of fractions that define the partitioning scheme to the `strata` object, which evaluates the Beale load and returns the associated RMSE estimate.

The genetic algorithm package, in turn, compares the current RMSE value with other previously calculated values (associated with other partitioning strategies), and 'evolves' a new partitioning strategy to test. This process is repeated hundreds or thousands of times in an attempt to find a optimum partitioning strategy.

```{r, genetic_algorithm}

 m <- mcga( popsize=200,
            chsize=4,
            minval=0.0,
            maxval=1.0,
            elitism=10,
            maxiter=40,
            crossprob=1.0,
            mutateprob=0.05,
            evalFunc=mystrata$update_rmse )

 cat("Best chromosome:\n")
 cat(m$population[1,],"\n")
 cat("Cost: ",m$costs[1],"\n")

# create a copy of the strata object
best_strata <- mystrata$clone()

# rerun with the optimized parameter
best_strata$partitions$update( sort( m$population[ 1, ]) )
best_strata$rearrange_stratums( best_strata$partitions$get_date_values() )

best_strata$calc_loads()
results_df <- best_strata$summarize_results()

# round values to produce a more readable table
results_df <- data.frame(lapply(results_df, function(y) if(is.numeric(y)) round(y, 2) else y)) 
```

The results of the load calculation with optimum stratification using 5 stratums now looks like this:

```{r, echo=FALSE, results='asis'}
knitr::kable( results_df, digits=2 )
```

## Original AutoBeale results for the same data file

For reference, here are the summarized results generated by Peter Richard's original AutoBeale code:

----------------------------------------------------------------------------

    Stratification applied:

    block  stratum   start      end   

    1       1     19970101 19970420
    2       2     19970421 19970526
    3       3     19970527 19970817
    4       4     19970818 19971201
    5       5     19971202 19971231


    Summary over 5 strata:
    
    Mean daily loading:      
                             459187.6  g/day
                          167603487.   g/year
                             167603.4  kg/year

    Mean square error:   893200710.529 ( g/day)**2
                         118996664.660 (kg/year)**2

    based on     9.348 degrees of freedom
    The 95% confidence interval half-width is       24539.3375 kg/year
-----------------------------------------------------------------------------
